Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models

### 2020-06-10 analysis/paper/paper.Rmd
edits to text in Biological Expectations from QRF Model section

---
title: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@merck.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@merck.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@merck.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
# csl: "../templates/canadian-journal-of-fisheries-and-aquatic-sciences.csl"
abstract: |
  Needs more...

@@

The results of the QRF parr capacity model for spring/summer Chinook salmon meet many biological expectations. Focusing on the partial dependence plots (Figure \@ref(fig:pdp-figure)), the QRF model predicts capacity to increase when the wetted width, discharge and the width-to-depth ratio grow, when temperatures are cooler, when there is less fine sediment, when there is more fish cover, when channel unit frequency increases and when the standard deviation of the wetted depth (a proxy for complexity) increases. These are all patterns that emerged from the fish-habitat data, and where available, those fish-habitat relationships match what has been found qualitatively in other studies (**need more citations??**) [@Mossop2006].

The biggest driver of capacity is stream size, whether measured by wetted width or discharge, which should be unsurprising since we are using fish per meter as our response. In many ways, these metrics define habitat quantity. However, other metrics help define habitat quality, such as cooler temperatures in August, less pool-tail fine sendiment, and higher channel unit frequencies (another way of saying more pool-riffle interfaces).

@@

>>>>

---
title: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@merck.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@merck.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@merck.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
# csl: "../templates/canadian-journal-of-fisheries-and-aquatic-sciences.csl"
abstract: |
  Needs more...

@@

The results of the QRF parr capacity model for spring/summer Chinook salmon meet many biological expectations. Focusing on the partial dependence plots (Figure \@ref(fig:pdp-figure)), the QRF model predicts capacity to increase when the wetted width, discharge and the width-to-depth ratio grow, when temperatures are cooler, when there is less fine sediment, when there is more fish cover, when channel unit frequency increases and when the standard deviation of the wetted depth (a proxy for streambed complexity) increases. These are all patterns that emerged from the fish-habitat data, and where available, match those fish-habitat relationships identified qualitatively in other studies [@Mossop2006].

The biggest driver of capacity identified in this study is stream size, whether measured by wetted width or discharge, which should be unsurprising since we are using fish per meter as our response. In many ways, these metrics define habitat quantity. However, other metrics used in our QRF model help define habitat quality, such as cooler temperatures in August, less pool-tail fine sendiment, and higher channel unit frequencies (a measure of habitat complexity and another way of saying more pool-riffle interfaces) and fish cover.

@@

Moreover, predictions of available habitat capacity seemed to pass the "eye test" when comparing predictions to the actual physical habitat at CHaMP survey locations. That is, areas with higher predicted capacities had habitat that, in our opinion, looked like habitat that would support more juvenile Chinook salmon, whereas areas with lower predicted capacities looked like poorer rearing habitat. As examples, Figure \@ref(fig:site-photo-high) is a photograph of a CHaMP survey site where the model predicted high capacity and is a site with an increases width-to-depth ratio and fish cover (in the form of mature willow and cottonwood riparian cover), contains pool-riffle sequences, and some variation in wetted depths. Alternatively, Figure \@ref(fig:site-photo-low) shows a site where the model predicted a much lower capacity, and visually, is much poorer rearing habitat for juvenile Chinook salmon e.g., lower wetted widths, discharge, width-to-depth ratio and standard deviation of depth. Pairing predictions of capacity from the QRF model with images and expert opinion of good versus bad habitat will be an important ongoing validation in addition to our comparisons with spawner-recruit models.
